---
title: "Class 3 - Digging into data"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

# Digging into Data
You'll all be working with your own data, so we'll review a concept with the `iris` data set, and then you'll have a few mintues to try it on your own. We'll cover the basic ways I explore data before diving in more depth. 

### Reading in data: two ways
```{r echo = T}
##---- for CSV files ----##
data <- read.csv("teams.csv", as.is=T)

##---- for XLSX files ----##
#install.packages("xslx")
library(xlsx)
data <- read.xlsx("NAME.xlsx", sheetIndex = 1) #you want the first (or only) sheet
  
## for the demo
iris_data <- iris

```
### Some mild data cleaning

Open up your data from the Evnironment pane - how does it look? Do you need to fix anything?
```{r}
####################################################################################
##--- Common issue #1: Column 1 is just row numbers, so let's remove it ---##
####################################################################################
iris_data <- iris
iris_data <- iris_data[, -1] # the - sign just removes that column number



####################################################################################
##--- Common issue #2: The column names get messed up (usually the first one), so let's rename --##
####################################################################################
##--- This is optional, but essential if you need to join data together on a key (advanced)
iris_data <- iris
names(iris_data)[1] <- "Sepal_Length_woohoo"



####################################################################################
##--- Common issue #3: Missing data is NA when it should be 0, or 0 when it should be NA ---##
####################################################################################
#install.packages("visdat")
#library(visdat)
visdat::vis_miss(iris)


iris_data <- iris
## replacing all entries that are 0 to be NA instead
iris_data$Sepal.Length[iris_data$Sepal.Length == 0] <- NA

## replacing all entries that are NA to be 0 instead
iris_data$Sepal.Length[is.na(iris_data$Sepal.Length)] <- 0



####################################################################################
##--- Common issue #4: (during analysis) oops, I deleted all my data! ----##
####################################################################################


iris_data <- iris_data[iris_data$Species == "Yeehaw", ] #oops!

## It's ok! Just re-run the lines where you read in the data!

```

## Data digging - let's go!

IMPORTANT: you need to keep track of interesting findings or weird results. Even something like a running page in a notebook. Lots of data analysis is exploration, so you need to ensure you're keeping track of what you've come upon. 

### Comparing means across groups
```{r}
## loading up the right libraries
library(dplyr)
library(ggplot2)

## I got bored with the flowers - let's use titanic data
#install.packages("titanic")
library(titanic)
titanic_data <- titanic_train

####################################################################################
##----Task 1: pick a group (strings or numbers) and summarize all other numeric columns - then dig into a finding! 
####################################################################################
summary(titanic_data)

titanic_data %>%
  group_by(Sex) %>%
  summarise_if(is.numeric, mean, na.rm=T)

## example of digging futher into a finding - men had fewer family members - what % were alone?
titanic_data %>%
  count(Sex, SibSp, Parch) # this isn't that useful :(

titanic_data %>%
  group_by(Sex) %>%
  summarize(PercentAlone = mean(SibSp == 0 & Parch == 0), 
            NumAlone = sum(SibSp == 0 & Parch == 0), 
            total = n()) # how many of each sex are there?

## save.csv()

####################################################################################
### Copy & paste & try it on your own!
####################################################################################

```

### Plotting
```{r}
library(dplyr)
library(ggplot2)
####################################################################################
## Task 2: Scatter plots to understand relationship between variables 
####################################################################################
iris_data <- iris
ggplot(data = iris_data, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() 

ggplot(data = iris_data, aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point()

ggplot(data = iris_data, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point()

## same thing but exploring the titanic data set
## adapted from here - https://www.kaggle.com/josepandreu/titanic-visualization-with-ggplot2
ggplot(data = titanic_data, aes(Pclass, Age, colour = factor(Survived))) +
    geom_jitter(size = 2) + 
    facet_grid(Sex ~ .) + 
    theme_light()

ggplot(data = titanic_data, aes(x= Fare, Age, colour = factor(Survived))) +
    geom_jitter(size = 2) + 
    theme_light()



####################################################################################
## Copy & paste & try on your own!
####################################################################################

```

### Making a new column
```{r}
####################################################################################
## Task 3 - make a new column that you want to use in your analysis
####################################################################################
## copying code from Class 2
iris_data <- iris
iris_data$SepalSize <- "Average"
iris_data$SepalSize[iris_data$Sepal.Length < 5.8 & iris_data$Sepal.Width < 3] <- "Small"
iris_data$SepalSize[iris_data$Sepal.Length > 5.8 & iris_data$Sepal.Width > 3] <- "Big"

## other example using titanic data
library(titanic)
titanic_data <- titanic_test

## how does age differ among those who were alone & those who weren't?
titanic_data$Alone <- 0
titanic_data$Alone[titanic_data$SibSp == 0 & titanic_data$Parch == 0] <- 1

titanic_data %>%
  group_by(Alone, Sex) %>%
  summarise(AvgAge = mean(Age, na.rm=T))


####################################################################################
## Make a new column & then either plot it or do some numeric summaries (recycle above code!)
####################################################################################

```

